---
title: "shim: A Simple HTTP Service for SciDB"
author: "B. W. Lewis <blewis@paradigm4.com>"
date: "11/9/2015"
output:
  html_document:
    theme: cerulean
    css: shimstyle.css
---


# What's new (for SciDB 15.7)

## SciDB authentication replaced PAM

SciDB version 15.7 introduced new native SciDB authentication. Shim has replaced PAM-based authentication with the new SciDB authentication methods. In order to maintain compatibility with pre-15.7 shim clients, the new SciDB authentication process mimics the PAM process using the `/login` service, see TLS Authentication for examples. At some point in the future, shim will no longer use the `/login` or `/logout` services.

## Use the new **/upload** service in place of **/upload_file**

The new `/upload` service is faster and simpler to use.

## Irritating trailing carriage return/newlines have been reduced

This is in particular the case for the `/new_session` service. You should no longer need to strip off trailing newlines from the response.

## New config settings for maximum concurrent shim sessions and SciDB instance

Shim can now use any SciDB instance as a query coordinator. That means that you can run more than one shim service, for example one service per SciDB cluster node.


# Overview

Shim is a web service that exposes a very simple API for clients to interact with SciDB over HTTP connections. The API consists of a small number of URIs (described in detail below), including: `/login`, `/new_session`, `/release_session`, `/execute_query`, `/cancel`, `/read_lines`, `/read_bytes`, `/upload_file`, `/version`, and `/logout`.

There are more direct ways to talk to SciDB over a network. See https://github.com/artyom-smirnov/scidb4py for a great example of direct network communication with SciDB using Python and Google protocol buffers (written by Artyom Smirnov, one of the core SciDB developers). The protocol buffers approach is comparatively low-level, but possibly more efficient than shim. Shim is designed to be a convenient and easy way to talk to SciDB. Note that both shim and Artyom's Python interface are community open source projects; they are not official components of SciDB.

Shim clients begin by requesting a session ID from the service, then running a query and releasing the session ID when done. Session IDs are distinct from SciDB query IDs--a session ID groups a SciDB query together with server resources for input and output to the client.


# Configuration

Shim runs as a system service or may be invoked directly from the command line. See the shim manual page for command-line options (type `man shim` from a terminal). Service configuration is determined by the `/var/lib/shim/conf` configuration file. The default conf file is a sample that displays the default configuration options, which are listed as one key=value pair per line. Available options include: 

```
#ports=8080,8083s
scidbport=1239
instance=0
tmp=/home/scidbdata/000/0/tmp
#user=root
#max_sessions=50
#timeout=60
```

Each option is described below.


## Ports and Network Interfaces

Shim listens on default ports 8080 (open, not encrypted), and 8083 (TLS encrypted) on all available network interfaces. Ports and listening interfaces are configured with the command line '-p' option or with the 'ports=' option in the `/var/lib/shim/conf` file when shim is run as a service. The ports/interface specification uses the following syntax:

```
[address:]port[s][,[address:]port[s]][,...]
```
where:

- *address* indicates an optional I.P. address associated with a specific available network device, only specify this if you want to restrict shim to operate on a specific network device.
- *port* indicates a required available port number
- *s* is an optional single character 's' suffix indicating that TLS/SSL should be used on that port.

Here are some examples of possible port configurations:

<table>
<tr><td>5555s <td style='padding-left:30px;'> <td> Listen only on port 5555 (TLS/SSL).
<tr><td>127.0.0.1:8080,1234s <td> <td>List on port 8080 but only on the local loopback interface; listen on port 1234(TLS/SSL) on all interfaces.
</table>

## SciDB Port

Shim runs on the same computer as a SciDB coordinator. Set the 'scidbport' option to select the coordinator database port to locally connect to. The default SciDB database port value is 1239 (see the SciDB configuration manual for more information). Since any SciDB instance can act as a query
coordinator, it is possible to configure multiple shim services, for example one shim service per computer IP address.

## Instance

Set the SciDB instance number to use as a query coordinator. Make sure that this option is set together with the corresponding SciDB port
number for the instance, and also set a corresponding temporary I/O location that the instance has read/write access to.

## Temporary I/O space and streaming

Shim's default behavior caches the output of SciDB queries in files on the SciDB server; set that file directory location with the config file tmp option or the command-line -t argument. This temporary director is also used to upload data from clients over the http connection for input into SciDB. Select a directory that is writable by the shim user (see the user option).

If you install shim from an RPM or Debian package as a service, the package will configure shim to use a SciDB data directory for temporary
storage. You can edit the config file and restart shim to change that default.


## User

The user that the shim service runs under (shim can run as a non-root user).


## Max sessions

Set the maximum number of concurrent _shim_ sessions, beyond which clients recieve an HTTP 'resource unavailable' error.

## Timeout

Set the time in seconds after which an _inactive_ session is considered timed out and a canidate for resource de-allocation.
After sessions time out their resources are not freed unless the need to be to satisfy additional session demands. See the
lazy timeout section below. Active sessions that are waiting on SciDB query results or transferring data are not subject
to timeout and may run indefinitely.



## TLS/SSL Certificate

Shim supports TLS/SSL encryption. Packaged versions of shim (RPM and Debian packages) generate a self-signed certificate and 4096-bit RSA key when shim is installed. The certificate is placed in `/var/lib/shim/ssl_cert.pem`. If you would prefer to use a different certificate, replace the automatically generated one.



# API Reference

Examples use the URL `http://localhost:8080` or `https://localhost:8083` (TLS) below. Parameters are required unless marked optional. All shim API services support CORS.

## Limits

HTTP 1.1 clients or greater are required.

All HTTP query parameters are passed to the service as string values. They are limited to a maximum of 4096 characters unless otherwise indicated (a notable exception is the SciDB query string parameter, limited to 262,144 characers).

HTTP query string parameters that represent numbers have limits. Unless otherwise indicated whole-number values (session ID, number of bytes to return, etc.) are interpreted by shim as signed 32-bit integers and are generally limited to values between zero and 2147483647. Values outside that range will result in an HTTP 400 error (invalid query).

## Response codes

Possible responses for each URI are listed below. HTTP status code 200 always indicates success; other standard HTTP status codes indicate various errors. The returned data may be UTF-8 or binary depending on the request and is always returned using the generic application/octet-stream MIME type. Depending on the request, data may used chunked HTTP transfer encoding and may also use gzip content encoding.


## Basic digest access authentication

Shim supports basic digest access authentication. (See https://en.wikipedia.org/wiki/Digest_access_authentication and the references therein for a good description of the method.) Enable digest access authentication by creating an .htpasswd file in shim's default `/var/lib/shim/wwwroot` directory (the .htpasswd file must be located in shim's wwwroot directory, which can be changed with the command line switch -r. The format of the file must be:

```
username1:password1
username2:password2
...
```

Use plain text passwords in the file, and consider changing the permissions of the file to restrict access.
Delete the .htpasswd file to disable basic digest access authentication.

Basic digest authentication works on plain or TLS-encrypted connections but can not be used in combination with SciDB authentication (see below).


## TLS/SSL encryption and SciDB authentication

Shim optionally exposes both open and encrypted (HTTPS/TLS) services. SciDB authentication requires encrypted HTTPS connections. SciDB authentication was introduced in SciDB version 15.7 and uses user names and passwords. The shim implementation uses the `/login` service in order to maintain compatibility with the PAM-based pre-15.7 shim API. Future versions of shim may not use the `/login` service.

SciDB authentication is simple to use:

- Contact the `/login` service over SSL with username and password query string parameters. The `/login` service returns an authentication token if its successful. The token is simply "username:password".
- Use all the other URI services normally, but append the auth=<token> query string parameter, replacing <token> with the response from the `/login` service above.
- Optionally log out of the system by contacting the `/logout` service with your authentication token.
- Shim's SciDB authentication mechanism only works over encrypted connections. Username and password query string parameters only appear in encrypted form over the network, and are never logged by the shim service log.

Here is a full example using the `wget` program at the command line talking to shim running on localhost, port 8083.

```{r, engine='bash', eval=FALSE}
# Log in with username=root password=paradigm4
# (assuming that such a user exists on the system that SciDB is running on):

wget -q --no-check-certificate -O - "https://localhost:8083/login?username=root&password=Paradigm4"
root:Paradigm4         # (authentication token returned on success)

# As a quick check, we'll try to obtain a new session without using the auth
# parameter (this will return an error):

wget --no-check-certificate -O - "https://localhost:8083/new_session"
HTTP request sent, awaiting response... 400 ERROR

# Now, let's try again with auth and run a simple query and this time get a new session ID:

wget -q --no-check-certificate -O - "https://localhost:8083/new_session?auth=root:Paradigm4"
0    # (our session ID)
wget -q --no-check-certificate -O - "https://localhost:8083/execute_query?id=0&query=list('functions')&save=dcsv&auth=root:Paradigm4"
wget -q --no-check-certificate -O - "https://localhost:8083/read_lines?id=0&n=0&auth=root:Paradigm4" | head -n 10
{No} name,profile,deterministic,library
{0} '%','double %(double,double)',true,'scidb'
{1} '%','int16 %(int16,int16)',true,'scidb'
{2} '%','int32 %(int32,int32)',true,'scidb'
{3} '%','int64 %(int64,int64)',true,'scidb'
{4} '%','int8 %(int8,int8)',true,'scidb'
{5} '%','uint16 %(uint16,uint16)',true,'scidb'
{6} '%','uint32 %(uint32,uint32)',true,'scidb'
{7} '%','uint64 %(uint64,uint64)',true,'scidb'
{8} '%','uint8 %(uint8,uint8)',true,'scidb'
{9} '*','double *(double,double)',true,'scidb'

wget -q --no-check-certificate -O - "https://localhost:8083/release_session?id=0&auth=root:Paradigm4"
```
Again, simply append your authentication token to your HTTP requests to use the authenticated service.


## Generic API Workflow

```
/new_session
/execute_query
/read_lines or /read_bytes
/release_session
```

## Generic Authenticated API Workflow

```
/login
/new_session
/execute_query
/read_lines or /read_bytes
/release_session
...
```


## Streaming and Compression

Specify `stream > 0` in the `/execute_query` service endpoint to stream SciDB results directly to the client using the HTTP 1.1 chunked transfer encoding mechanism. With this option, a named pipe is created in the temporary I/O directory instead of a file. See the `/execute_query` API doc below for more information. This option is automatically set by the compression option below.

Specify `compression=n`, for values of `n` in `1,...,9` in `/excute_query` to stream gzip compressed (RFC 1952) results directly to the client using HTTP 1.1 chunked transfer encoding. The stream compression level is set by the compression query parameter in the range `1,...,9` (1 fast but poor compression to 9 slow but great compression). It's up to the client to decompress the data. Setting this option autmatically also sets the stream option to a nonzero value of 2 indicating gzip-compressed streaming.



## API Service Endpoints

The R examples below use the `httr` package. We try to illustrate API calls with real examples using either curl or R.
See https://github.com/Paradigm4/shim/tree/master/tests for additional examples.

### **/version**
<table>
<tr><td>DESCRIPTION  <td>Print the shim code version string
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS
<tr><td>RESPONSE <td>**Success** HTTP 200 and text version string value in text/plain payload 
<tr><td>EXAMPLE (curl)
<td>
```{r, engine='bash', comment=NA}
curl -f -s http://localhost:8080/version 
```
<tr><td>EXAMPLE (R)
<td>
```{r, comment=NA}
httr::GET("http://localhost:8080/version")
```
</table>


### **/new_session**
<table>
<tr><td>DESCRIPTION  <td>Request a new HTTP session from the service.
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS <td>
- **auth** _optional_ authentication token (required for authenticated connections). 
<tr><td>RESPONSE <td>
- Success: HTTP 200 and text session ID value in text/plain payload 
- Failure (out of resources/server unavailable): HTTP 503 
- Invalid request (encrypted only -- this means auth is missing): HTTP 400 
- Not authorized (encrypted only): HTTP 401
<tr><td>EXAMPLE (curl)
<td>
```{r, engine='bash', comment=NA}
curl -f -s http://localhost:8080/new_session 
```
<tr><td>EXAMPLE (R)
<td>
```{r, comment=NA}
id = httr::GET("http://localhost:8080/new_session")

(id = rawToChar(id$content))
```
</table>

### **/release_session**
<table>
<tr><td>DESCRIPTION  <td>Release an HTTP session.
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS <td>
- **id** an HTTP session ID obtained from `/new_session`
- **auth** _optional_ authentication token (required for authenticated connections). 
<tr><td>RESPONSE <td>
- Success: HTTP 200 
- Failure (Session not found): HTTP 404 
- Failure (invalid http query): HTTP 400 
- Not authorized (encrypted only): HTTP 401
<tr><td>EXAMPLE (R)
<td>
```{r, comment=NA}
httr::GET(sprintf("http://localhost:8080/release_session?id=%s",id))
```
</table>


### **/execute_query**
<table>
<tr><td>DESCRIPTION  <td>Execute a SciDB AFL query.
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS <td>
- **id** an HTTP session ID obtained from `/new_session`
- **auth** _optional_ authentication token (required for authenticated connections). 
- **query** AFL query string, encoded for use in URL as required, limited to a maximum of 1,000,000 characters
- **save** optional SciDB save format string, limited to a maximum of 4096 characters; Save the query output in the specified format for subsequent download by `read_lines` or `read_bytes`. If the save parameter is not specified, don't save the query output. 
- **release** optional 0 or 1: if 1 then release_session as soon as query completes. The default value is 0 if not specified (see additional notes below).
- **stream** optional 0, 1 or 2: if 1 then stream query result; if 2 then stream a gzip compressed (RFC 1952) query result with an unspecified (automatic) compression level; otherwise send query result through a server-side output file (the default). steam>0 also sets release=1.
- **compression** optional 0 to 9: Sets the steaming compression level from 0 (no compression), 1 (fast/light compression) to 9 (slow/high compression). Setting this option automatically also sets the stream option to stream=2.
<tr><td>RESPONSE <td>
- Success: HTTP 200 text/plain (SciDB Query ID)
- Failure (SciDB not available error): HTTP 503 text/plain (ERROR TEXT) 
- Failure (SciDB query error): HTTP 500 text/plain (SCIDB ERROR TEXT) 
- Failure (out of memory error): HTTP 507 text/plain (SCIDB ERROR TEXT) 
- Failure (Invalid session): HTTP 404 
- Failure (invalid http query): HTTP 400 
- Not authorized (encrypted only): HTTP 401 
<tr><td>EXAMPLE (curl)
<tr><td>NOTES <td>
Shim only supports AFL queries. 
<p>
Remember to URL-encode the SciDB query string.
<p>
500 and 503 errors result in removal of the web session ID and related resources (thus, `release_session` does not have to be called after such an error). 
<p>
This method blocks until the query completes unless stream>0. When stream>0, this method returns immediately with the query ID (before the query completes). 
<p>
Do not specify the option release=1 when stream=0 and when the save option is also set, or output will not be available to `read_bytes` or `read_lines`. Instead, explicitly call `release_session` after reading is complete. 
<p>
Setting stream > 0 is an experimental feature added after version 14.7. Setting stream > 0 sends query results through a server-side named pipe into the client without storing the query results on the server. This option uses the HTTP 1.1 chunked transfer encoding; stream termination is indicated by a chunk of length zero. 
<p>
When stream>0 is set then release=1 is also automatically set. That means that clients are free to not call `/release_session` after receiving their stream data via `/read_lines` or `/read_bytes`;however, calling `/release_session` will not cause problems and clients can still do it if they wish. 
<p>
NOTE: WHEN stream>0, SciDB QUERIES REMAIN ACTIVE UNTIL THE CLIENT RECEIVES THE RESULT DATA. IT'S UP TO THE CLIENT TO RECEIVE THE DATA TO TERMINATE THE QUERY. 
<p>
Set stream=2 to enable gzip-compressed streaming (RFC 1952). When stream=2 one may also set the compression level with the compression option in the range 0 to 9. If the compression query parameter is specified, then stream is automatically set to 2.
<tr><td>EXAMPLE (R)
<td>
```{r, comment=NA}
# Obtain a shim session ID
id = httr::GET("http://localhost:8080/new_session")
session = rawToChar(id$content)

# Construct the query request
query = sprintf("http://localhost:8080/execute_query?id=%s&query=consume(list())&release=1",
                session)
ans = httr::GET(query)

# The response in this example is just the SciDB query ID:
(rawToChar(ans$content))


# Release our session ID
httr::GET(sprintf("http://localhost:8080/release_session?id=%s", session))
```
<tr><td>EXAMPLE w/ERROR (R)
<td>
```{r, comment=NA}
id = httr::GET("http://localhost:8080/new_session")
session = rawToChar(id$content)
query = sprintf("http://localhost:8080/execute_query?id=%s&query=consume(42)&release=1",
                session)
httr::GET(query)
httr::GET(sprintf("http://localhost:8080/release_session?id=%s", session))
```
<tr><td><td>See `/read_lines` and `/read_bytes` below for running queries that return results and downloading them.
</table>

### **/cancel**
<table>
<tr><td>DESCRIPTION  <td>Cancel a SciDB query associated with a shim session ID.
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS <td>
- **id** an HTTP session ID obtained from `/new_session`
- **auth** _optional_ authentication token (required for authenticated connections). 
<tr><td>RESPONSE <td>
- Success: HTTP 200 
- Failure (session not found): HTTP 404 
- Failure (invalid http query): HTTP 400 
- Failure (could not connect to SciDB): HTTP 503 
- Not authorized (encrypted only): HTTP 401
<tr><td>EXAMPLE (R)
<td>
```{r, comment=NA, eval=FALSE}
# An example cancellation of a query associated with shim ID 19 (not run)
httr::GET("http://localhost:8080/cancel?id=19")
```
</table>


### **/read_lines**
<table>
<tr><td>DESCRIPTION  <td>Read text lines from a query that saves its output.
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS <td>
- **id** an HTTP session ID obtained from `/new_session`
- **n**  the maximum number of lines to read and return between 0 and 2147483647.
- **auth** _optional_ authentication token (required for authenticated connections). 
<tr><td>RESPONSE <td>
- Success: HTTP 200 followed by application/octet-stream query result (up to n lines) 
- Failure (invalid HTTP query string): HTTP 400 
- Failure (session not found): HTTP 404 
- Failure (end of file): HTTP 410 
- Failure (invalid request): HTTP 414 
- Failure (SciDB server error): HTTP 500
- Failure (could not connect to SciDB server error): HTTP 503
- Failure (server out of memory): HTTP 507
- Not authorized (encrypted only): HTTP 401
<tr><td>NOTES <td>
Set n=0 to download the entire output buffer. The parameter n is automatically set to zero when streaming is indicated. Simply terminate the read early if desired in those cases (see example above). You should almost always set n=0.
<p>
Be sure to properly url-encode special characters like the plus sign (+) in the request.
<p>
When n>0, iterative requests to `read_lines` are allowed, and will return at most the next n lines of output. Use the 410 error code to detect end of file output. Don't use this option if you can avoid it.
<p>
Note that query results are _always_ returned as application/octet-stream.
<tr><td>EXAMPLE (curl)
<td>
```{r, comment=NA, engine='bash'}
s=`curl -s "http://localhost:8080/new_session"`
curl -s "http://localhost:8080/execute_query?id=${s}&query=list('functions')&save=dcsv"
curl -s "http://localhost:8080/read_lines?id=${s}&n=10"
curl -s "http://localhost:8080/release_session?id=${s}"
```
<tr><td>EXAMPLE streaming (curl) <td>
```{r, comment=NA, engine='bash'}
s=`curl -s "http://localhost:8080/new_session"`
curl -s "http://localhost:8080/execute_query?id=${s}&query=list('functions')&save=dcsv&stream=1"
curl -s "http://localhost:8080/read_lines?id=${s}" | head -n 10
curl -s "http://localhost:8080/release_session?id=${s}"
```
<tr><td>EXAMPLE compressed streaming (curl) <td>
```{r, comment=NA, engine='bash'}
s=`curl -s "http://localhost:8080/new_session"`
curl -f -s -k "http://localhost:8080/execute_query?id=${s}&query=list('functions')&save=dcsv&compression=9" >/dev/null || fail
curl -f -s -k "http://localhost:8080/read_bytes?id=${s}&n=0"  | zcat | head -n 10
curl "http://localhost:8080/release_session?id=${s}"
```
</table>


### **/read_bytes**
<table>
<tr><td>DESCRIPTION  <td>Read bytes lines from a query that saves its output.
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS <td>
- **id** an HTTP session ID obtained from `/new_session`
- **n**  the maximum number of bytes to read and return between 0 and 2147483647.
- **auth** _optional_ authentication token (required for authenticated connections). 
<tr><td>RESPONSE <td>
- Success: HTTP 200 followed by application/octet-stream query result (up to n lines) 
- Failure (invalid HTTP query string): HTTP 400 
- Failure (session not found): HTTP 404 
- Failure (end of file): HTTP 410 
- Failure (invalid request): HTTP 414 
- Failure (SciDB server error): HTTP 500
- Failure (could not connect to SciDB server error): HTTP 503
- Failure (server out of memory): HTTP 507
- Not authorized (encrypted only): HTTP 401
<tr><td>NOTES <td>
Set n=0 to download the entire output buffer. The parameter n is automatically set to zero when streaming is indicated. Simply terminate the read early if desired in those cases (see example above). You should almost always set n=0.
<p>
Be sure to properly url-encode special characters like the plus sign (+) in the request.
<p>
When n>0, iterative requests to `read_bytes` are allowed, and will return at most the next n lines of output. Use the 410 error code to detect end of file output. Don't use this option if you can avoid it.
<p>
Note that query results are _always_ returned as application/octet-stream.
<tr><td>EXAMPLE (curl)
<td>
```{r, comment=NA, engine='bash'}
# Obtain a new shim session ID
s=`curl -s "http://localhost:8080/new_session"`

# The URL-encoded SciDB query in the next line is just:
# build(<x:double>[i=1:10,10,0],u)
curl -s "http://localhost:8080/execute_query?id=${s}&query=build(%3Cx:double%3E%5Bi=1:10,10,0%5D,i)&save=(double)"

# Pass the double-precision binary result through the `od` program to view:
curl -s "http://localhost:8080/read_bytes?id=${s}" | od -t f8

# Release the session
curl -s "http://localhost:8080/release_session?id=${s}"
```
</table>


### **/upload_file**
<table>
<tr><td>DESCRIPTION  <td>Upload a file to the HTTP service using a multipart/file POST method.
<tr><td>METHOD <td>POST
<tr><td>PARAMETERS <td>
- **id** an HTTP session ID obtained from `/new_session`
- A valid multipart/file POST body -- see the example below
- **auth** _optional_ authentication token (required for authenticated connections). 
<tr><td>RESPONSE <td>
- Success: HTTP 200 and the name of the file uploaded to the server in a text/plain response. 
- Failure (invalid HTTP query string): HTTP 400 
- Failure (Session not found): HTTP 404 
- Failure (Server error): HTTP 500 
- Not authorized (encrypted only): HTTP 401 
<tr><td>NOTES <td>Try to avoid using this method. It's fairly slow to transfer data and difficult to get the POST
body message right. Instead use the faster and simpler `/upload` method shown below.
<tr><td>EXAMPLE (curl)
<td>
```{r, engine='bash'}
# Upload 5 MB of random bytes
id=$(curl -s  "http://localhost:8080/new_session")
dd if=/dev/urandom bs=1M count=5  | \
  curl -s --form "fileupload=@-;filename=data" "http://localhost:8080/upload_file?id=${id}"
curl -s "http://localhost:8080/release_session?id=${id}"
```
</table>


### **/upload**
<table>
<tr><td>DESCRIPTION  <td>Upload data to the HTTP service using a basic POST method.
<tr><td>METHOD <td>POST
<tr><td>PARAMETERS <td>
- **id** an HTTP session ID obtained from `/new_session`
- A valid POST body -- see the example below
- **auth** _optional_ authentication token (required for authenticated connections). 
<tr><td>RESPONSE <td>
- Success: HTTP 200 and the name of the file uploaded to the server in a text/plain response. 
- Failure (invalid HTTP query string): HTTP 400 
- Failure (Session not found): HTTP 404 
- Failure (Server error): HTTP 500 
- Not authorized (encrypted only): HTTP 401 
<tr><td>NOTES <td>Use the returned server-side file name in later calls, for example to `execute_query`.
<p>
This method is faster and easier to use than the older `/upload_file` method.
<tr><td>EXAMPLE (curl)
<td>
```{r, engine='bash'}
id=$(curl -s "http://localhost:8080/new_session")

# Upload 5 MB of random bytes
dd if=/dev/urandom bs=1M count=5  | \
  curl -s --data-binary @- "http://localhost:8080/upload?id=${id}"

curl -s "http://localhost:8080/release_session?id=${id}"
```
<tr><td>EXAMPLE (R) <td>
```{r}
# Obtain a shim session ID
id = httr::GET("http://localhost:8080/new_session")
session = rawToChar(id$content)

# Upload a character string:
httr::POST(sprintf("http://localhost:8080/upload?id=%s", session), body="Hello shim")

# Release our session ID
httr::GET(sprintf("http://localhost:8080/release_session?id=%s", session))
```
</table>

### **/login**
<table>
<tr><td>DESCRIPTION  <td>Authenticate a SciDB user.
<tr><td>METHOD <td>GET
<tr><td>PARAMETERS <td>
- **username** SciDB username
- **password** SciDB password. 
<tr><td>RESPONSE <td>
- Success: HTTP 200 and a text authentication token to use with the auth parameter in subsequent HTTP queries. 
- Failure (invalid http query): HTTP 400
- Not authorized: HTTP 401
<tr><td>NOTE<td>**/login requires a TLS-encrypted connection.** It won't work over un-encrypted shim connections.
<p>The returned token is simply username:password. This method is only present for backwards-compatibility and
will be removed in the future. You don't need to use it.
<tr><td>EXAMPLE<td>

```{}
# Example service URL for user 'scidb' and password 'paradigm4':
https://localhost:8083/login?username=scidb&password=paradigm4 

# Example response:
HTTP/1.0 200 OK
Content-Length: 11
Content-Type: text/plain

scidb:paradigm4
```

</table>


# Orphaned Sessions

Shim limits the number of simultaneous open sessions. Absent-minded or malicious clients are prevented from opening too many new sessions repeatedly without closing them (which could eventually result in denial of service). Shim uses a lazy timeout mechanism to detect unused sessions and reclaim them. It works like this:

- The session time value is set to the current time when an API event finishes.
- If a new_session request fails to find any available session slots, it inspects the existing session time values for all the sessions, computing the difference between current time and the time value. If a session time difference exceeds a timeout value, then that session is harvested and returned as a new session.
- Operations that may take an indeterminate amount of time like file uploads or execution of SciDB queries are protected from harvesting until the associated operation completes.

The above scheme is called lazy as sessions are only harvested when a new session request is unable to be satisfied. Until that event occurs, sessions are free to last indefinitely.

Shim does not protect against uploading gigantic files nor from running many long-running SciDB queries. The service may become unavailable if too many query and/or upload operations are in flight; an HTTP 503 (Service Unavailable) error code is returned in that case.


Copyright (C) 2015, Paradigm4.